---
title: "Class Prediction Analysis for Gene Expression Data"
author: "Beatriz Manso"
date: '2022-04-20'
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
    code_folding: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For each sample that is provided with this package, it becomes possible
to predict which class it belongs to using a multivariate predictor. In
addition to Compound Covariate Predictor, Diagonal Linear Discriminant
Analysis, Nearest Neighbor Predictor, Nearest Centroid Predictor, and
Support Vector Machine Predictor, there are many other multivariate
classification methods. A graphical representation of how accurately
this multivariate class predictor can determine all class predictions
requested is provided in this package.

The complete evaluation of the process is carried out by using
cross-validation methods including leave-one-out cross-validation,
k-fold validation, and bootstrap validation with 0.632+. Each
classifier's performance is examined along with a cross-validated
estimate of misclassification rate. The full dataset can be used to
classify new samples based on specific predictors and multivariate
predictors.

# Methods

Set working directory:

```{r}
setwd("C:/Users/manso/OneDrive - University of West London/MSc Bioinformatics - UWL/6.BGA - Bioinformatics and Genome Analysis/week 5 - Microarray analysis/practical")
```

Install packages and load libraries:

```{r}
if (!require("BiocManager", quietly = TRUE))
 install.packages("BiocManager")
BiocManager::install("ROC")

#Class prediction package
install.packages("https://brb.nci.nih.gov/BRB-ArrayTools/RPackagesAndManuals/classpredict_0.2.tar.gz", repos = NULL, type="source")

library(classpredict)


```

## 1. Get Built-in sample data - Cancer data

Expression data:

```{r}
dataset <- "Brca"

x <- read.delim(system.file("extdata", paste0(dataset, "_LOGRAT.TXT"), package = "classpredict"), header = FALSE)

head(x)
```

## 2. Class Information

```{r}
expdesign <- read.delim(system.file("extdata", paste0(dataset, "_EXPDESIGN.txt"), package = "classpredict"), as.is = TRUE)

head(expdesign)

```

## 3. Class Prediction Analysis

The "classPredict" function calculates multiple classifiers that are
used to predict the class of a new sample, implementing the class
prediction tool with multiple methods in BRB-ArrayTools. This package
provides test.classPrediction for a quick start of class prediction
analysis over one of the built-in sample data (i.e., "Brca", "Perou",
and "Pomeroy").

```{r}
res1 <- test.classPredict('Brca', outputName = "ClassPrediction_Brca", 
generateHTML = TRUE)

## Getting analysis results ...
res2 <- test.classPredict('Pomeroy', outputName = "ClassPrediction_Pomeroy", 
generateHTML = TRUE)

## Getting analysis results ...
res3 <- test.classPredict('Perou', outputName = "ClassPrediction_Brca", 
generateHTML = TRUE)


```

## 4. List Objects In The Results

```{r}
names(res1)
names(res2)
names(res3)
```

### Explanation about each object:

**res\$performClass** is a data frame with the performance of
classifiers during cross-validation:

```{r}
res1$performClass[1:11,]
```

**res\$percentCorrectClass** is a data frame with the mean percent of
correct classification for each sample using different prediction
methods:

```{r}
res1$percentCorrectClass
```

-   "ccp": Compound Covariate Predictor
-   "bcc": Bayesian Compound Covariate Predictor
-   "dlda": Diagonal Linear Discriminant Analysis
-   "knn": 1-Nearest Neighbor/ 3-Nearest Neighbor
-   "nc": Nearest Centroid
-   "svm": Support Vector Machine

**res\$predNewSamples** is a data frame with predicted class for each
new sample. NC means that a sample is not classified. In this example,
there are four new samples:

```{r}
res1$predNewSamples[1:4,]

```

**res\$probNew** is a data frame with the predicted probability of each
new sample belonging to the class (BRCA1) from the the Bayesian Compound
Covariate method:

```{r}
res1$probNew[1:4,]
```

Note:

|                              |                                                                                                                                                                                                                                                                                                                                                                             |
|:----------------------------:|:----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
|   **res\$classifierTable**   | Data frame with composition of classifiers such as geometric means of values in each class, p-values and Gene IDs                                                                                                                                                                                                                                                           |
|     **res\$probInClass**     | Data frame with predicted probability of each training sample belonging to a class during cross-validation from the Bayesian Compound Covariate                                                                                                                                                                                                                             |
|     **res\$CCPSenSpec**      | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the Compound Covariate Predictor Classifier                                                                                                                                                                                                           |
|     **res\$LDASenSpec**      | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the Diagonal Linear Discriminant Analysis Classifier.                                                                                                                                                                                                 |
|     **res\$K1NNSenSpec**     | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the 1-Nearest Neighbor Classifier                                                                                                                                                                                                                     |
|     **res\$K3NNSenSpec**     | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the 3-Nearest Neighbor Classifier                                                                                                                                                                                                                     |
|   **res\$CentroidSenSpec**   | Data frame with performance (i.e., sensitivity, \* specificity, positive prediction value, negative \| prediction value) of the Nearest Centroid Classifier                                                                                                                                                                                                                 |
|     **res\$SVMSenSpec**      | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the Support Vector Machine Classifier                                                                                                                                                                                                                 |
|     **res\$BCPPSenSpec**     | Data frame with performance (i.e., sensitivity, specificity, positive prediction value, negative prediction value) of the Bayesian Compound Covariate Classifier                                                                                                                                                                                                            |
|  **res\$weightLinearPred**   | Data frame with gene weights for linear predictors such \*\* as Compound Covariate Predictor, Diagonal Linear Discriminant Analysis and Support Vector Machine                                                                                                                                                                                                              |
| **res\$thresholdLinearPred** | Contains the thresholds for the linear prediction rules related with res\$weightLinearPred. Each prediction rule is defined by the inner sum of the weights (wiwi) and log expression values (xixi) of significant genes. In this case, a sample is classified to the class BRCA1 if the sum is greater than the threshold; that is, ∑iwixi\>threshold∑iwixi\>threshold |
|     **res\$GRPCentroid**     | Data frame with centroid of each class for each predictor gene                                                                                                                                                                                                                                                                                                              |
|       **res\$pmethod**       | Vector of prediction methods that are specified                                                                                                                                                                                                                                                                                                                             |
|      **res\$workPath**       | Path for Fortran and other intermediate outputs                                                                                                                                                                                                                                                                                                                             |

## 5. Producing ROC Curves

Cross-validation ROC curves are provided for Compound Covariate
Predictor, Diagonal Linear Discriminant Analysis and Bayesian Compound
Covariate Classifiers.

```{r}
plotROCCurve(res1,"ccp")

```

```{r}
plotROCCurve(res1,"dlda")
```

```{r}
plotROCCurve(res1,"bcc")
```
